---
output:
  word_document: default
  html_document: default
---

```{r}
# Greeshma Chowdary Peddineni
```

```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
library(readr)
library(forcats)
```



```{r}
new_now_covid_22 <- readRDS("C:/Users/Dell/Downloads/covid22.rds")
```

Let's find how many different vaccine manufacturers are there?
```{r}
uniq_manufacturers <- unique(new_now_covid_22$VAX_MANU)
num_manufacturers <- length(uniq_manufacturers)

cat("There are", num_manufacturers, "different COVID-19 vaccine manufacturers:\n")
cat(paste("", uniq_manufacturers), sep = "\t")

```

Create a plot to show adverse events by month for each manufacturer
```{r}
new_now_covid_22 <- new_now_covid_22 %>%
  mutate(RECVDATE = as.Date(RECVDATE, format="%Y-%m-%d"),
         Month = floor_date(RECVDATE, "month"))  

monthly_events <- new_now_covid_22 %>%
  group_by(Month, VAX_MANU) %>%
  summarise(Event_Count = n(), .groups = 'drop')

ggplot(monthly_events, aes(x = Month, y = Event_Count, color = VAX_MANU)) +
  geom_line(size = 2) +  
  labs(title = "Monthly Adverse Events Reported by month for each Vaccine Manufacturer",
       x = "Reported Month",
       y = "Number of Adverse Events",
       color = "Vaccine Manufacturer") +
  theme_light() +  
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +  
  scale_color_manual(values = c("MODERNA" = "orange",
                                "PFIZER\\BIONTECH" = "yellow",   
                                "JANSSEN" = "green",
                                "NOVAVAX" = "red",
                                "UNKNOWN MANUFACTURER" = "violet"),
                     na.translate = FALSE) +  
  theme(plot.title = element_text(face = "bold", size = 10),  
        axis.text.x = element_text(angle = 45, hjust = 1)) 
```

Let's see which state had the highest number of events?
```{r}
new_now_covid_22 <- new_now_covid_22 %>%
  mutate(
    STATE = as.character(STATE),  
    STATE = trimws(STATE)         
  )

invisible(unique(new_now_covid_22$STATE))

state_event_counts <- new_now_covid_22 %>%
  filter(!is.na(STATE) & STATE != "") %>% 
  group_by(STATE) %>%
  summarise(Event_Count = n(), .groups = 'drop') %>%
  arrange(desc(Event_Count))  

highest_state <- state_event_counts[1, ]

cat("The state with the highest number of reported events is", highest_state$STATE, 
    "with", highest_state$Event_Count, "events.\n")
```

For Idaho, Texas, and Michigan, show number of events by gender?
```{r}
selected_states <- c("TX", "MI", "ID")
filtered_data <- new_now_covid_22 %>% filter(STATE %in% selected_states)


gender_counts <- filtered_data %>%
  group_by(STATE, SEX) %>%
  summarise(Event_Count = n(), .groups = 'drop')

gender_counts
```


Let's find how many ER/ED visits were reported?
```{r}
er_ed_visits_count <- new_now_covid_22 %>%
  filter(!is.na(ER_ED_VISIT) & ER_ED_VISIT != "") %>%
  nrow()

cat("Total ER/ED visits reported:", er_ed_visits_count, "\n")
```

Which manufacturer is associated with the lowest number of reported ER/ED visits?
```{r}
manufacturer_er_ed_counts <- new_now_covid_22 %>%
  filter(ER_ED_VISIT == TRUE | ER_ED_VISIT == "Y") %>%
  group_by(VAX_MANU) %>%
  summarise(ER_ED_Visit_Count = n(), .groups = 'drop')


low_manufacturer <- manufacturer_er_ed_counts %>%
  arrange(ER_ED_Visit_Count) %>%
  slice(1)

cat(low_manufacturer$VAX_MANU, "manufacturer is associated with the lowest number of reported ER/ED with", low_manufacturer$ER_ED_Visit_Count, "visits.")
```



